Multi-level generalizations of collective spin dynamics and squeezing

### 2019-09-27 collective_sun/sun_notes.tex
fixed minor error in definition of Dicke states

The drive operators have the convenient feature that $D_{1,0}$ and
$D_{1,1}$ are (up to a gauge transformation) respectively proportional
to the Hamiltonians induced on an $n$-level spin by an external
magnetic field and a classical driving laser; loosely speaking, one
should identify $\p{D_{1,0},D_{1,1}}\sim\p{S_\z,S_\x}$, where
$S_\alpha$ is a collective SU(2) spin-$\alpha$
operator\cite{perlin2019shorttime}.  More concretely, we can embed the
Hilbert space of a single $n$-level spin into the
permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
2-level spins via
\begin{align}
  \ket{\mu}_{\t{single}} \to \ket{\mu+I}_{\t{collective}}
  \propto S_+ \ket\dn^{\otimes\p{n-1}},
\end{align}
where $I\equiv\p{n-1}/2$ is a total spin,
$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
2-level spins.  This embedding takes
\begin{align}
  \p{D_{1,0},D_{1,1},D_{1,-1}}
  \to \p{S_\z,S_\x,S_\y}
  \times \sqrt{\f{3}{I \p{I+1/2} \p{I+1}}}.
\end{align}

@@

An external driving laser addressing a spin-$1/2$ particle on a
lattice will induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \f12 \Omega \sum_j
  \p{e^{i\phi_j} c_{j,\up}^\dag c_{j,\dn}
    + e^{-i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a drive strength, $j$ indexes a lattice site, and
$\phi_j$ is a site-dependent phase.  The multi-level analogue of
\eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\f12 \Omega \sum_j \p{e^{i\phi_j} T_{1,1}^{(j)}
    + e^{-i\phi_j} {T_{1,1}^{(j)}}^\dag}.
  \label{eq:H_drive}
\end{align}
We can make this drive spatially homogenous via the gauge
transformation $c_{j\mu} \to e^{i\mu\phi_j} c_{j\mu}$, after which the
drive becomes
\begin{align}
  H_{\t{drive}}
  = -\f12 \Omega \sum_j\p{T_{1,1}^{(j)} + {T_{1,1}^{(j)}}^\dag}
  = \f12 \Omega \sum_j D_{1,1}
  = \Omega \D_{1,1},
\end{align}
where $\D_{LM}\equiv\sum_jD_{LM}^{(j)}/2$ is a collective drive
operator, with e.g.~$\D_{1,1}=S_\x$ in the case of $n=2$.  The SU($n$)
symmetry of collective interactions implies that the interaction
Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is invariant under
this gauge transformation, but the single-particle Hamiltonian in
\eqref{eq:H_lat} transforms as
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat_SOC}
\end{align}
where $\phi$ is the relative phase of the driving laser between
adjacent lattice sites.  In the basis of drive operators, this
Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)},
  \label{eq:H_lat_drive}
\end{align}
with effective driving field strengths
\begin{align}
  B_L\p{q,\phi} = \p{-1}^L f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu;I\mu,L,0} f_L\p{\mu\phi},
  &&
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}

>>>>

The drive operators have the convenient feature that $D_{1,0}$ and
$D_{1,1}$ are (up to a gauge transformation) respectively proportional
to the Hamiltonians induced on an $n$-level spin by an external
magnetic field and a classical driving laser; loosely speaking, one
can identify $\p{D_{1,0},D_{1,1}}\sim\p{S_\z,S_\x}$, where $S_\alpha$
is a collective SU(2) spin-$\alpha$
operator\cite{perlin2019shorttime}.  More concretely, we can embed the
Hilbert space of a single $n$-level spin into the
permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
2-level spins via
\begin{align}
  \ket{\mu}_{\t{single}} \to \ket{I+\mu}_{\t{collective}}
  \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
\end{align}
where $I\equiv\p{n-1}/2$ is a total spin,
$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
2-level spins.  This embedding takes
\begin{align}
  \p{D_{1,0},D_{1,1},D_{1,-1}}
  \to \p{S_\z,S_\x,S_\y}
  \times \sqrt{\f{3}{I \p{I+1/2} \p{I+1}}}.
\end{align}

@@

An external driving laser addressing a spin-$1/2$ fermion on a lattice
will induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \f12 \Omega \sum_j
  \p{e^{i\phi_j} c_{j,\up}^\dag c_{j,\dn}
    + e^{-i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a drive strength, $j$ indexes a lattice site, and
$\phi_j$ is a site-dependent phase.  The multi-level analogue of
\eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\f12 \Omega \sum_j \p{e^{i\phi_j} T_{1,1}^{(j)}
    + e^{-i\phi_j} {T_{1,1}^{(j)}}^\dag}.
  \label{eq:H_drive}
\end{align}
We can make this drive spatially homogenous via the gauge
transformation $c_{j\mu} \to e^{i\mu\phi_j} c_{j\mu}$, after which the
drive becomes
\begin{align}
  H_{\t{drive}}
  = -\f12 \Omega \sum_j\p{T_{1,1}^{(j)} + {T_{1,1}^{(j)}}^\dag}
  = \f12 \Omega \sum_j D_{1,1}
  = \Omega \D_{1,1},
\end{align}
where $\D_{LM}\equiv\sum_jD_{LM}^{(j)}/2$ is a collective drive
operator.  The SU($n$) symmetry of collective interactions implies
that the interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int_drive} is invariant under this gauge transformation,
but the single-particle Hamiltonian in \eqref{eq:H_lat} transforms as
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat_SOC}
\end{align}
where $\phi$ is the relative phase of the driving laser between
adjacent lattice sites.  In the basis of drive operators, this
Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)},
  \label{eq:H_lat_drive}
\end{align}
with effective driving field strengths
\begin{align}
  B_L\p{q,\phi} = \p{-1}^L f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu;I\mu,L,0} f_L\p{\mu\phi},
  &&
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}
